\chapter{Authors and Contributors}%
\label{cha:contributors}
\index{Authors}
\index{Contributors}

\textbf{Adam Williams} - wrote the original complete Cinelerra package Copyrighted under GPL and continues to release new versions yearly; this is the HV version <broadcast@earthling.net> .

Sam - designed and engineered the \CGG{} website and continuously maintains and updates the Operating system and applications.

Willam Morrow - was responsible for the current \CGG{} software version which emphasized crash-free stability; he started adding fixes and features in 2007, merged his version based on the HV version with the CV version in 2016, continued to add new features and fixes and incorporate desirable updates from HV and CV until November 2020.
 
Software developers working on \CGG{} between the years of 2021 to 2023 along with some areas where they have been crucial (in alphabetical order by first name):
\begin{itemize}
 \item Andrew Randrianasulu - continuous ongoing support; bug fixes; library upgrades with relevant patches; some nice enhancements; created and debugged builds: NetBSD and the Android tablet using Termux.
 \item Andras Reuss - created very useful Autosave backup feature in Preferences with Andrew assist.
 \item Georgy Salnikov - added all of the Help software in the program allowing access to HTML manual pages; major improvement in the Motion plugin; provided scripts and program mods to convert Latex manual to HTML.
 \item Manuel Virgil - DPX list writer developer; EXR package upgrade to 3.1.4 (not currently in use).
 \item Mat Nieuwenhoven - creator of the AppImage build procedure; LV2 specific control patch; maintains \CGG{} wikipaedia entry.
\end{itemize} 

Personnel providing current recent testing, examples, demos, and work on the manual for the \CGG{} version (in alphabetical order by first name):
\begin{itemize}
 \item Andrea Paz - main person who keeps the Manual updated; does builds for early testing of changes; provides examples and demos.
 \item Andrey Spitsyn - responsible for the huge task of getting the \CGG{} Manual in LaTex.
 \item DeJay - expert user and frequent responder to users in the Forum.
 \item Glen MacArthur - provides A/V Linux distro which includes \CGG{}.
 \item Pierre Autourduglobe - designed and thoroughly tested Interview Mode for implementation; major tester of Mixers; provided reason for getting the Shuttle to work; tested other new features.
 \item Phyllis Smith - facilitator of checking into GIT the developers and contributors changes to the Programs and to the Manual; tester; creates monthly AppImages when needed.
 \item RafaMar - updated Spanish translations; instrumental in getting friend's help in adding the Help button to the Batch Render menu; provided additional ShapeWipe transitions.
 \item Terje Hanssen - especially tests Bluray video media creation providing invaluable feedback.
\item \textbf{+ many others}, some of who wish to remain anonymous.
\end{itemize}

Software Developers in the earlier years of \CGG{} (alphabetical by last name):
\begin{itemize}
\item Balakhnin, Alexander - author of the Descratch plugin.
\item Messick, Eric - provided the initial Shuttle code (FixedImagePhoto.com/Contact).
\item Olaf - Cakewalk and Neophyte theme developer - Cakewalk is the default theme due to its being modern; tested and provided much feedback.
\item Roenitz, Frederic - wrote the VP8/VP9 render formats.
\end{itemize}

Contributors and Software Developers through the years either on HV or CV, the Community Version.  They are listed by last name in
 alphabetical order, along with some of their specific expert work - often just a single known item listed, and it may not even have been the most important when they obviously did so much more.  The CV developers were
responsible for many improvements based on the HV version to include bugfixing, creating build systems,
internationalization work, plugins, and IIRC ffmpeg support appeared in one of the CV forks.
\begin{itemize}
\item Arendt, David - early tester providing significant feedback.
\item Basto, EaSergio M. - completely replaced/fixed autogen.sh; removed old outdated fonts.
\item Baverstock, Richard - assisted with working on the SVG plugin; added tape controls in the capture monitor window; major early tester and contributor.
\item Bielefeldt, Karl - created gradient wipe plugin; worked on outdated mpeg2enc.
\item Brosius, Kevin - provided many patches, for example fixed scroll bar resizing in bclistbox.C; major early contributor.
\item Cornet, Jerome - updated Deinterlace plugin for temporal field swap.
\item Coulon, Jean-Luc - updated French translation; provided patches for translation gettextization strings.
\item Delannoy, Florent - C41 plugin to convert a 16 bit C41 digital intermediate negative file to positive image.
\item Diniz, Rafael - worked on the Brazillian Portuguese translation; created Slackware packages.
\item Dumuid, Pierre - assisted in merge efforts; added column in File Load dialog box for sorting by extension; wrote Selective Temporal Averaging plugin. 
\item Erdelyi, Georgely - early on provided patches for libsndfile, titler, and other programs to compile on Suse.
\item Ferralis, Nicola - updated Ubuntu PPA over many years; along with Slock Ruddy, responsible for GreyCStoration plugin (currently inactive in \CGG{}), worked on Italian translation.
\item Ferrer, Alex - originally put up a Cinelerra manual TWiki; maintained the wiki.
\item Flaming, Benjamin - created the Diffkey plugin. 
\item Geddes, Tyler - contributed to CV Version.
\item Harang, Camille - worked on fixing guicast programs for big Endian; part of small team of 3 that used Cinelerra to make short documentaries/interviews.
\item Heen, Tollef Fog - contributed to CV Version.
\item Iniguez, Gustavo - contributed to HV version.
\item Joyeux, Sylvain - worked on the UI; worked on configuration/build scripts.
\item Kielb, Andreas - created Firewire I/O patches.
\item Konink, Stefan de - added patch for playing without sound.
\item Kurz, Nathan - added a YUV4MPEG pipe output dialog to the user interface; yuvstream.C program developer; early contributor.
\item Maufrais, Nicolas - collaborated on the Cinelerra-CV manual.
\item Mekkes, Greg - added patch to Motion Blur plugin.
\item Menk, Michael Eric - contributed to CV Version.
\item Montgomery, Monty - Blue Banana color modification plugin; fileffmpeg developer in CV fork.
\item Mueller, Johannes - author of keyboard tracks up/down shortcut and horizontal scrolling with the mouse wheel.
\item Muylkens, Koen - created the Blue Dot theme.
\item Pesegov, Valentin - added Russian translation.
\item Poure, Jean-Michel  - provided fixes in build procedures for Debian; first entry in Mailing List September 2004.
\item Rampino, Paolo - aka Akirad, created several colorful/artistic themes; developer of the cinecutie fork (now inactive); worked on Italian translation.
\item Reinholdtsen, Petter - assisted in early hardware upgrade; core developer; moved over new features from HV to CV; fixed many, many bugs.
\item Robak, Herman - created initial draft repositories; major early contributor.
\item Runkaru, Einar - numerous bug fixes; for a long time maintaned a separate branch and currently has his own personal CVE fork; noted first entry in Mailing List in January 2010.
\item Schmarsel, Danny - past website designer and administrator for Cinelerra-CV website; created Cinelerra LiveDVD for Slackware (now inactive); provided German translation.
\item Seigne, Eric - initial French translation.
\item Setti, Riccardo - contributed to CV Version.
\item \textbf{Sixt, Johannes - long time contributor from early on and continuous}; notable for merging into Cinelerra-CV the new releases of HV.
\item Stewart, Joe - worked on Inkscape for Cinelerra; created Motion Blur plugin.
\item Streetman, Dan - developed patch to correct text\_to\_seconds; added patch to Titler for timecode functionality; patches to ReframeRT.
\item Taraba, Mark - provided support for all types of PNGs; first Mailing List entry in August 2004.
\item Thaeter, Christian -  maintained server to host Cinelerra-CV website.
\item Tori, Andraz - responsible for Cinelerra x86 Debian packages; worked on an SVG plugin; added overlaying speedup patch; major early contributor and lots of debugging.
\item Traniello, Raffaella - best known for "Cinelerra for Grandma" still active at \url{http://g-raffa.eu/Cinelerra/HOWTO/basics.html} .
\item Vladimirsky, Igor - was igor\_ubuntu at one time; developer of the Russian po file; currently supports the Cinelerra-CV site; and much, much more work; noted first entry in Mailing List 2014.
\item Vosseler, Hermann - developer of the Reroute plugin used to transfer the Alpha
channel and/or the RGB/YUV Components from source track to target track; significant
Bezier automation patch for fades, camera and projector.
\item Wulff, Jonas - patched Histogram plugin to provide a split output screen.
\newline
\item \textbf{+ many others}.
\end{itemize}
